---
title: ""
author: ""
date: "2023-03-15"
format:
  bookup-html:
    embed-fonts: false
toc: true
---

## Chapter 03: Discrete Random Variables


## Problem 3.1

**Let $X$ be a discrete random variable with probability mass function given by**

$$
p(x) = \begin{cases}
\frac{1}{4} &  x = 0 \\
\frac{1}{2} &  x = 1 \\
\frac{1}{8} &  x = 2 \\
\frac{1}{8} &  x = 3 \\
\end{cases}
$$

a. Verify that $p(x)$ is a probability mass function.
b. Find $P(X \geq 2)$.
c. Find $P(X \geq 2 | X \geq 1)$.
d. Find $P(X \geq 1 \cup X \geq 2)$.


### Part (a)

To be a valid probability function $p(x)$ must satisfy the following conditions:

- $p(x) \geq 0$ for all $x$.
- $\sum_{x} p(x) = 1$.

The first condition is satisfied by construction. The second condition is satisfied by the following calculation:

$$
\sum_{x} p(x) = \frac{1}{4} + \frac{1}{2} + \frac{1}{8} + \frac{1}{8} = 1
$$

### Part (b)

$$
P(X \geq 2) = P(X = 2) + P(X = 3)
$$

$$
P(X \geq 2) = \frac{1}{8} + \frac{1}{8} = \frac{1}{4}
$$

### Part(c)

We can use conditional probability as:

$$
P(X \geq 2 | X \geq 1) = \frac{P(X \geq 2 \cap X \geq 1)}{P(X \geq 1)}
$$

Since $X \geq 1$ is a subset of $X \geq 2$, we have that $P(X \geq 2 \cap X \geq 1) = P(X \geq 2)$.

$$
P(X \geq 2 | X \geq 1) = \frac{P(X \geq 2)}{P(X \geq 1)}
$$

$$
P(X \geq 2 | X \geq 1) = \frac{\frac{1}{4}}{\frac{3}{4}} = \frac{1}{3}
$$

### Part (d)

$$
P(X \geq 1 \cup X \geq 2) = P(X \geq 1) + P(X \geq 2) \\ - P(X \geq 1 \cap X \geq 2)
$$

Again, Since $X \geq 1$ is a subset of $X \geq 2$, we have that $P(X \geq 2 \cap X \geq 1) = P(X \geq 2)$.

$$
P(X \geq 1 \cap X \geq 2) = (X \geq 2)
$$

$$
P(X \geq 1 \cap X \geq 2) = \frac{1}{8} + \frac{1}{8} = \frac{1}{4}
$$

And finally:

$$
P(X \geq 1 \cup X \geq 2) = \frac{3}{4} + \frac{1}{4} - \frac{1}{4} = \frac{3}{4}
$$

Wich make sense, since $P(X \geq 1 \cup X \geq 2) = P(X \geq 1)$

## Problem 3.2

**Let $X$ be a discrete random variable with probability mass function given by**

$$
p(x) = \begin{cases}
\frac{1}{4} &  x = 0 \\
\frac{1}{2} &  x = 1 \\
\frac{1}{8} &  x = 2 \\
\frac{1}{8} &  x = 3 \\
\end{cases}
$$

a. Use `sample` to create a sample of size 10000 from $X$ and *estimate* $P(X = 1) from your sample. Your result should be close to 1/2.
b. Use `table` on your sample from part (a) to *estimate* the pmf of $X$ from your sample. Your result should be similar to the pmf given in the problem.


### Part (a)

```{r}
x <- sample(0:3, size = 10000, replace = TRUE, prob = c(1/4, 1/2, 1/8, 1/8))
mean(x == 1)
```

### Part (b)

```{r}
x <- sample(0:3, size = 10000, replace = TRUE, prob = c(1/4, 1/2, 1/8, 1/8))
table(x) / 10000
```

## Problem 3.3

**Let $X$ be a discrete random variable with probability mass function given by**

$$
p(x) = \begin{cases}
\frac{C}{4} &  x = 0 \\
\frac{C}{2} &  x = 1 \\
C           &  x = 2 \\
0           & \text{otherwise}
\end{cases}
$$

Find the value of $C$ that makes $p(x)$ is a probability mass function.

$$
\sum_{x} p(x) = \frac{C}{4} + \frac{C}{2} + C = 1
$$

$$
C = \frac{4}{7}
$$

## Problem 3.4

**Give an example of a probability mass function $p$ whose associated random variable has mean 0.**

By definition, the mean of a random variable is:

$$
\mu = \sum_{x} x p(x)
$$

We know the pmf is positive and the sum of all probabilities is 1, so we can choose any symetric random variable over the real numbers, with equal probabilities for both sides.

$$
p(x) = \begin{cases}
\frac{1}{4} &  x = -2 \\
\frac{1}{4} &  x = -1 \\
\frac{1}{4} &  x = 1 \\
\frac{1}{4} &  x = 2 \\
0           & \text{otherwise}
\end{cases}
$$

The mean is:

$$
\mu =  -2 \frac{1}{4} + -1 \frac{1}{4} + 1 \frac{1}{4} + 2 \frac{1}{4} = 0
$$

## Problem 3.5

**Find the mean of the random variable $X$ given in Exercise 3.1.**

$$
\mu = \sum_{x} x p(x) = 0 \frac{1}{4} + 1 \frac{1}{2} + 2 \frac{1}{8} + 3 \frac{1}{8} = \frac{9}{8}
$$

## Problem 3.6

Let $X$ be a random variable with pmf given by $p(x) = 1/10$ for $x = 1, \ldots, 10$ and $p(x) = 0$ for all other values of $x$. Find $E[X]$ and confirm your answer using simulation.

$$
E[x] = \sum_{x} x p(x) = 1 \frac{1}{10} + 2 \frac{1}{10} + \ldots + 10 \frac{1}{10} = 5.5
$$

```{r}
x <- sample(1:10, size = 10000, replace = TRUE, prob = rep(1/10, 10))
mean(x)
```

## Problem 3.7

**Suppose you roll two ordinary dice. Calculate the expected value of their product.**

First we need to find the pmf of the product of two dice:

| Product | Outcomes                       | Probability |
|---------|--------------------------------|-------------|
| 1       | (1, 1)                         | 1/36        |
| 2       | (1, 2), (2, 1)                 | 2/36        |
| 3       | (1, 3), (3, 1)                 | 2/36        |
| 4       | (1, 4), (4, 1), (2, 2)         | 3/36        |
| 5       | (1, 5), (5, 1)                 | 2/36        |
| 6       | (1, 6), (6, 1), (2, 3), (3, 2) | 4/36        |
| 8       | (2, 4), (4, 2)                 | 2/36        |
| 9       | (3, 3)                         | 1/36        |
| 10      | (2, 5), (5, 2)                 | 2/36        |
| 12      | (3, 4), (4, 3), (6, 2), (2, 6) | 4/36        |
| 15      | (3, 5), (5, 3)                 | 2/36        |
| 16      | (4, 4)                         | 1/36        |
| 18      | (6, 3), (3, 6)                 | 2/36        |
| 20      | (5, 4), (4, 5)                 | 2/36        |
| 24      | (6, 4), (4, 6)                 | 2/36        |
| 25      | (5, 5)                         | 1/36        |
| 30      | (6, 5), (5, 6)                 | 2/36        |
| 36      | (6, 6)                         | 1/36        |

And then we can calculate the expected value:

$$
E[x]  = 1 \frac{1}{36} + 2 \frac{2}{36} + \ldots + 36 \frac{1}{36} = 12.25
$$

We can also use simulation to confirm our answer:

```{r}
dice_1 <- sample(1:6, size = 10000, replace = TRUE)
dice_2 <- sample(1:6, size = 10000, replace = TRUE)
mean(dice_1 * dice_2)
```
